On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-04-03 manuscripts/PNAS/chazthm_pnas.tex
update but still TODOs

Our main technical finding is that a sufficient diversity of measurements generated from any dictionary satisfying the spark condition 
has a stable sparse representation.  More specifically, such dictionaries are uniquely identifiable from as few as \mbox{$N = m(k-1){m \choose k} + m$} sparse linear combinations of their columns up to an error that is linear in the measurement noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $(n, m, k)$ satisfy CS inequality \eqref{CScondition}, in almost all cases the dictionary learning problem is well-posed given enough data (Cor.~\ref{ProbabilisticCor}). Importantly, these guarantees hold without assuming the recovered matrix satisfies the spark condition. Our results  also apply when only an upper bound on the number of dictionary elements is known. The explicit, algorithm-independent criteria we provide should be a useful tool in the theory of sparse dictionary learning.  % (Thm.~\ref{DeterministicUniquenessTheorem2})

@@

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{\hat m}$ with  \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ must have $\hat m \geq m$ and, provided $p = \hat m - \ell(\hat m - m)$ is positive, necessarily satisfy:
\begin{align}\label{Cstable}
\|(\mathbf{A}_J- \mathbf{B}_{\hat J} \mathbf{PD} )_j\|_2 \leq C_1 \varepsilon \ \  \text{for all $j \in [J]$},
\end{align}
for some $J \in {[m] \choose p}$, $\hat J \in {[\hat m] \choose p}$, permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$.

@@

\begin{corollary}
Fix $n, \ell$, and $k < m$. Let $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfy the spark condition \eqref{SparkCondition} and $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} have at least \mbox{$\left[ (k-1){ m \choose k} + 1 \right] + (k-1)^2{m \choose k-1}$} vectors supported in general linear position on each set in an $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP. Given any $\delta_1, \delta_2 \geq 0$ and $\varepsilon \leq \varepsilon(\delta_1, \delta_2)$,  % (\eqref{epsdel}).
% $\varepsilon < L_2(\mathbf{A}) / C_1$.
%There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.
every solution $\mathbf{B}, \mathbf{\hat x}_i, \ldots, \mathbf{\hat x}_N$ to the $\ell_0$ optimization problem:
\begin{align}\label{minsum}
\arg \min \sum_{i = 1}^N \|\mathbf{\hat x}_i\|_0 \ \ \text{s.t.} \ \ \mathbf{B}\mathbf{\hat x}_i = \mathbf{A}\mathbf{x}_i, \ \ \text{$i \in [N]$}, %\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon, \ \ \text{$i \in [N]$},
\end{align}
necessarily satisfies the recovery inequalities in \eqref{def1}. % \eqref{Cstable} and \eqref{b-PDa}.
\end{corollary}

\begin{proof}
We derive a lower bound on the number of $k$-sparse $\mathbf{\hat x}_i$ and then apply Thm.~\ref{DeterministicUniquenessCorollary}. 
First, note that $\|\mathbf{\hat x}_i\|_0 \neq 0$ for all $i$, since otherwise we would have $\mathbf{A}\mathbf{x}_i = \mathbf{B}\mathbf{\hat x}_i = \mathbf{0} \implies \mathbf{x}_i = \mathbf{0}$ by the spark condition, contradicting the general linear position of the $\mathbf{x}_i$. Let $n_p$ be the number of $\mathbf{\hat x}_i$ with $\|\mathbf{\hat x}_i\|_0 = p$.  %, so that $\sum_{p = 1}^{m} n_p = N$. 
Since the $\mathbf{x}_i$ are all $k$-sparse, by \eqref{minsum} we have:
\begin{align}
k \sum_{p = 1}^{m} n_p = kN \geq \sum_{i=1}^N \|\mathbf{x}_i\|_0 \geq \sum_{i=1}^N \|\mathbf{\hat x}_i\|_0 = \sum_{p=1}^{m} p n_p.
\end{align}
%\[\implies \sum_{p = 1}^k (k-p)n_p \geq \sum_{p = k+1}^{m} (p-k) n_p.\]
Hence,
\begin{align}
\sum_{p = k+1}^m n_p \leq \sum_{p = k+1}^m (p-k) n_p \leq \sum_{p = 1}^k (k-p)n_p \leq (k-1) \sum_{p = 1}^{k-1} n_p.
\end{align}
%\[ \implies \sum_{p = k+1}^m n_p \leq (k-1)^2 {\hat m \choose k-1} \]

By the spark condition and general linear position of the $\mathbf{x}_i$, every $k$ vectors $\mathbf{Ax}_i$ span a $k$-dimensional space; hence, at most $k-1$ vectors $\mathbf{\hat x}_i$ can share a support of size $k-1$ or less, and the number of $(k-1)$-sparse vectors $\mathbf{\hat x}_i$ is then bounded from above by $(k-1) { m \choose k-1}$. It follows that at least $N - (k-1)^2 {m \choose k-1}$ of the vectors $\mathbf{\hat  x}_i$ are $k$-sparse, and that there must then be some $N' \subseteq [N]$ for which for each $S \in E$ there are $(k-1){m \choose k}+1$ of the vectors $\mathbf{x}_i$ with $i \in [N']$ supported on $S$ and for which $\mathbf{\hat x}_i$ is $k$-sparse for all $i \in [N']$.  The result now follows directly from Thm.~\ref{DeterministicUniquenessCorollary}. % \eqref{Cstable} and \eqref{b-PDa} follow.
\end{proof}

@@

We  now show that $\pi$ is injective (and thus is a permutation if $\hat m = m$). Suppose that $\pi(i) = \pi(j) = \ell$ for some $i \neq j$ and $\ell$. Then, $\|c_{j}\mathbf{A}_{j} - \hat{c}_{j}\mathbf{B}_{\ell}\|_2 \leq \varepsilon$ and $\|c_{i}\mathbf{A}_{i} - \hat{c}_{i} \mathbf{B}_{\ell}\|_2  \leq \varepsilon$. Scaling and summing these inequalities by $|\hat{c}_{i}|$ and $|\hat{c}_{j}|$, respectively, and applying the triangle inequality, we obtain:
\begin{align*}%\label{contra}
(|\hat{c}_{i}| + |\hat{c}_{j}|) \varepsilon
&\geq\|\mathbf{A}(\hat{c}_{i}c_{j} \mathbf{e}_{j} - \hat{c}_{j}c_{i}\mathbf{e}_{i})\|_2 \nonumber \\ 
&\geq  \left( |\hat{c}_{i}| + |\hat{c}_{j}| \right) L_2(\mathbf{A}) \min_{\ell \in [m]} |c_\ell |,
\end{align*}
%
which contradicts the bound $\varepsilon < L_2(\mathbf{A})/C_1$. Hence, the map $\pi$ is injective. Setting $J = \pi([m])$ and letting $P = \left( \mathbf{e}_{\pi(1)} \cdots \mathbf{e}_{\pi(m)}\right)$ and $D = \text{diag}(\frac{\hat{c}_1}{c_1},\ldots,\frac{\hat{c}_m}{c_m})$, we see that \eqref{1D} becomes, for all $j \in [m]$:
\begin{align*}%\label{k=1result}
\|\mathbf{A}_j - (\mathbf{B}_J\mathbf{PD})_j\|_2 
= \|\mathbf{A}_j - \frac{\hat{c}_j}{c_j}\mathbf{B}_{\pi(j)}\|_2 
\leq \frac{\varepsilon}{|c_j|} 
\leq C_1\varepsilon.
\end{align*}
\end{proof}

@@

If for some  $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ and map $\pi: E \mapsto {[\hat m] \choose k}$ we have:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon \ \   \text{for all $S \in E$},
\end{align}
%
then $\hat m \geq m$ and, provided $p = \hat m - \ell(\hat m - m)$ is positive:
\begin{align}\label{MainLemmaBPD}
\|(\mathbf{A}_J - \mathbf{B}_{\hat J}\mathbf{PD})_j\|_2 \leq C_2 \varepsilon, \indent \text{for } j \in [J].
\end{align}
for some $J \in {[m] \choose p}$, $\hat J \in {[\hat m] \choose p}$, permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$.
\end{lemma}

@@

What we have shown here is that as an inverse problem to be approximated by a numerical algorithm, the sparse linear coding model  generally produces \textit{well-posed} optimizations.  This early concept of Hadamard \cite{Hadamard1902} can be paraphrased as the idea that good parameterized models should pose optimization problems with solutions that are stable; that is, continuous in the data.  In other words, perturbing a dataset by a small amount should only slightly influence the identifiability or uniqueness of parameters in the model, rather than discontinuously affect the character of optimization involved.  In this regard, we demonstrate a linear relationship between noise in the data and the unique solution in  sparse linear coding, with explicit constants expressed in terms of original latent parameters. 
Nonetheless, it would be of practical utility to determine the best possible dependence of $\varepsilon$ on $\delta_1, \delta_2$ in Def.\ref{maindef} as well as the minimal requirements on the number and diversity of generating codes, and we hope that other researchers continue to improve and extend our results.
% , although they are too computationally expensive to be of practical value.  We hope .

@@

A technical difficulty in our main theorem is the absence of a spark condition assumption on alternate sparse data factorizations.
Although mathematically interesting that no such requirement is necessary, there are other reasons to seek out such a theoretical guarantee. For instance, it is difficult to ensure an algorithm maintain a dictionary satisfying \eqref{SparkCondition} at each learning stage; indeed, even certifying  a dictionary has this property is likely intractable given its NP-hardness \cite{tillmann2014computational}.

In fact, though, uniqueness guarantees with minimal assumptions apply to all areas of data science and engineering that utilize learned sparse representations.  For example, several groups have utilized compressed sensing for signal processing tasks: MRI analysis \cite{lustig2008compressed},  image compression \cite{Duarte08}, and, more recently, the design of an ultrafast camera \cite{Gao14}. Given such effective uses of CS, it is only a matter of time before these systems incorporate sparse representational learning to encode and process data (e.g., in a device that learns structure from motion \cite{kong2016prior}).  In these cases, assurances such as those offered by our theorems ensure that different devices (with different initialization, samples, etc.) are equivalent as soon as enough data originate from a statistically identical system.
%Data science:
%  
%in a common application of compressed sensing in engineering

%\textbf{Theoretical Neuroscience}.
Additionally, in the field of theoretical neuroscience, sparse dictionary learning and related methods have recovered characteristic components of natural images \cite{Olshausen96, hurri1996image, bell1997independent, van1998independent} and sounds \cite{bellsejnowski1996, smithlewicki2006, Carlson12}, reproducing response properties of cortical neurons. Our results suggest that this correspondence could be due to  sparse representations being ``universal'' in natural data, an early mathematical idea in neural theory \cite{pitts1947}. Furthermore, our guarantees justify the hypothesis of \cite{Coulter10, Isely10} that sparse codes passed through a communication bottleneck in the brain can be recovered from random projections via (unsupervised) biologically plausible sparse dictionary learning (e.g., \cite{rehnsommer2007, rozell2007neurally, hu2014hebbian}).

@@

\begin{lemma}\label{NonEmptyLemma} 
Fix integers $\ell, m$ and $\hat m$. Suppose $E \subseteq 2^{[m]}$ is $\ell$-regular and satisfies the singleton intersection property. If there exists a map $\pi: E \to 2^{[\hat m]}$ such that $\sum_{S \in E} |S| = \sum_{S \in E} |\pi(S)|$ and:
\begin{align}\label{cond}
|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for all } F \subseteq E,
\end{align}
%
then $\hat m \geq m$ and the association $i \mapsto \cap \pi(F(i))$ defines an injective map from $[J]$ to $[\hat m]$ for some $J \subseteq [m]$ of size $\hat m - \ell(\hat m - m)$. In particular, if $\hat m = m$ then the map $\pi$ is induced by a permutation.
%$\pi(E)$ is $\ell$-regular and satisfies the singleton intersection property. In particular, $|\cap_{S \in F} S| = 1$ if and only if $|\cap_{S \in F} \pi(S)| = 1$. 
\end{lemma}
\begin{proof}
We first show that $\hat m \geq m$. Fix $T = \{(j, S): j \in \pi(S), S \in E\}$. Then $|T| = \sum_{S \in E} |\pi(S)| = \sum_{S \in E} |S| = \sum_{i \in [m]} \deg(i) = \ell m$ by regularity of $E$. If $\hat m < m$ then by pigeonholing the $m \ell \geq \ell (\hat m + 1) \geq [(\ell + 1) - 1] \hat m + 1$ elements of $T$ with respect to the set $[\hat m]$ of possible first indices, we see that there must exist at least $\ell + 1$ elements of $T$ having the same first index. But by \eqref{cond} there can be no more than $\ell$ elements of $T$ with a given first index, since $E$ is $\ell$-regular. Hence $\hat m \geq m$.

>>>>

Our main technical finding is that a sufficient quantity of measurements generated from any dictionary satisfying the spark condition 
has a stable sparse representation.  More specifically, such dictionaries are uniquely identifiable from as few as \mbox{$N = m(k-1){m \choose k} + m$} sparse linear combinations of their columns up to an error that is linear in the measurement noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $(n, m, k)$ satisfy CS inequality \eqref{CScondition}, in almost all cases the dictionary learning problem is well-posed given enough data (Cor.~\ref{ProbabilisticCor}). Importantly, these guarantees hold without assuming the recovered matrix satisfies the spark condition. Our results  also apply when only an upper bound on the number of dictionary elements is known. The explicit, algorithm-independent criteria we provide should be a useful tool in the theory of sparse dictionary learning.  % (Thm.~\ref{DeterministicUniquenessTheorem2})

@@

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{\hat m}$ with  \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ must have $\hat m \geq m$ and, provided $p = \hat m - \ell(\hat m - m)$ is positive, necessarily satisfy:
\begin{align}\label{Cstable}
\|(\mathbf{A}_J- \mathbf{B}_{\hat J} \mathbf{PD} )_j\|_2 \leq C_1 \varepsilon, \ \  \text{for all $j \in J$},
\end{align}
for some $J \in {[m] \choose p}$, $\hat J \in {[\hat m] \choose p}$, permutation matrix $\mathbf{P}$, and invertible diagonal matrix $\mathbf{D}$.

@@

\begin{corollary}
Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} for some $k < m$ and that $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ contain at least $\left[ (k-1){m \choose k} + 1 \right] + |E|k(k-1){m \choose k-1}$ vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP.  Then there exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ and set of vectors $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{\hat m}$ that solve:
\begin{align}\label{minsum}
\min \sum_{i = 1}^N \|\mathbf{\hat x}_i\|_0 \ \ 
\text{subject to} \ \ \|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon, \ \ \text{for all $i$},
\end{align}
%
must have $\hat m \geq m$ and, provided $p = \hat m - \ell(\hat m - m)$ is positive, necessarily satisfy:
\begin{align}\label{Cstable}
\|(\mathbf{A}_J- \mathbf{B}_{\hat J} \mathbf{PD} )_j\|_2 \leq C_1 \varepsilon \ \  \text{for all $j \in J$},
\end{align}
for some $J \in {[m] \choose p}$, $\hat J \in {[\hat m] \choose p}$, permutation matrix $\mathbf{P}$, and invertible diagonal matrix $\mathbf{D}$.
\end{corollary}

\begin{proof}
We derive a lower bound on the number of $k$-sparse $\mathbf{\hat x}_i$ and then apply Thm.~\ref{DeterministicUniquenessCorollary}. 
Let $n_p$ be the number of $\mathbf{\hat x}_i$ with $\|\mathbf{\hat x}_i\|_0 = p$.  %, so that $\sum_{p = 1}^{m} n_p = N$. 
Since the $\mathbf{x}_i$ are all $k$-sparse, by \eqref{minsum} we have:
\begin{align}
k \sum_{p = 0}^{m} n_p = kN \geq \sum_{i=0}^N \|\mathbf{x}_i\|_0 \geq \sum_{i=0}^N \|\mathbf{\hat x}_i\|_0 = \sum_{p=0}^{m} p n_p.
\end{align}
Hence,
\begin{align}\label{eqn}
\sum_{p = k+1}^m n_p \leq \sum_{p = k+1}^m (p-k) n_p \leq \sum_{p = 0}^k (k-p)n_p \leq k \sum_{p = 0}^{k-1} n_p.
\end{align}

Let $\mathbf{X}$ and $\mathbf{\hat{X}}$ be the $m \times N$ matrices with columns $\mathbf{x}_i$ and $\mathbf{\hat x}_i$, respectively, and let $I(S) = \{i \in [N]: \text{supp}(\mathbf{x}_i) \subseteq S\}$. It follows from the general linear position of the $\mathbf{x}_i$ and the linear independence of every $k$ columns of $\mathbf{A}$ that $L_k(\mathbf{AX}_{I(S)}) > 0$ for all $S \in E$; that is, for every $S \in E$, the columns of every $n \times k$ submatrix of $\mathbf{AX}_{I(S)}$ form a basis for $\text{\rm Span}\{\mathbf{A}_S\}$. Suppose there are $(k-1)|E| + 1$ vectors $\mathbf{x}_i$ for which the corresponding $\mathbf{\hat x}_i$ $(i \in J)$ share a support $\hat S$. By the pigeonhole principle, there must be some $S \in E$ supporting $k$ or more of these vectors $\mathbf{x}_i$; let them be indexed by $J \subseteq I(S)$. Fixing $\mathbf{y} \in \text{\rm Span}\{ \mathbf{A}_S \}$, there exists a unique $\mathbf{c} = (c_1, \ldots, c_k) \in \mathbb{R}^k$ such that $\mathbf{y} = \mathbf{AX}_{J}\mathbf{c}$. Setting \mbox{$\mathbf{\hat{y}} = \mathbf{B\hat{X}}_{J}\mathbf{c} \in \text{\rm Span}\{\mathbf{B}_{\hat S}\}$}, we have:
\begin{align*}
\|\mathbf{y} - \mathbf{\hat{y}}\|_2 
&= \|\sum_{i=1}^k c_i(\mathbf{AX}_J - \mathbf{B\hat{X}}_J)_i\|_2
\leq \varepsilon \sum_{i=1}^k |c_i| \\
&\leq \varepsilon \sqrt{k}  \|\mathbf{c}\|_2 
\leq \frac{\varepsilon}{L_k(\mathbf{AX}_J)} \|\mathbf{y}\|_2,
\end{align*}
where the last inequality is by definition of $L_k$. From Def.~\ref{dDef}, it follows that:
\begin{align}\label{rhs222}
d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{\hat S}\}) 
\leq \frac{\varepsilon}{ \min_{S \in E}L_k(\mathbf{AX}_{I(S)}) }.
\end{align}

Since (by our upper bound on $\varepsilon$) the RHS is strictly less than one, we have by \eqref{dimlem} that $k = \dim(\text{Span}(\mathbf{AX}_J)) \leq \dim(\text{Span}\{\mathbf{B}_{\hat S}\}) \leq |\hat S|$. Hence, if $|E|(k-1) + 1$ or more vectors $\mathbf{\hat x}_i$ share the same support, they cannot be $(k-1)$-sparse and the number of $(k-1)$-sparse vectors $\mathbf{\hat x}_i$ must then be bounded from above by $|E|(k-1){ m \choose k-1}$. Taking \eqref{eqn} into account, it follows that at least $N - |E|k(k-1){ m \choose k-1}$ of the vectors $\mathbf{\hat  x}_i$ are $k$-sparse, and there must then be some $N' \subseteq [N]$ for which for each $S \in E$ there are $(k-1){m \choose k}+1$ of the vectors $\mathbf{x}_i$ with $i \in [N']$ supported on $S$ and for which $\mathbf{\hat x}_i$ is $k$-sparse for all $i \in [N']$. The result now follows directly from Thm.~\ref{DeterministicUniquenessCorollary}.
\end{proof}

%\begin{corollary}
%Fix $n, \ell$, and $k < m$. Let $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfy the spark condition \eqref{SparkCondition} and $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} have at least \mbox{$\left[ (k-1){ m \choose k} + 1 \right] + (k-1)^2{m \choose k-1}$} vectors supported in general linear position on each set in an $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP. Given any $\delta_1, \delta_2 \geq 0$ and $\varepsilon \leq \varepsilon(\delta_1, \delta_2)$,
%every solution $\mathbf{B}, \mathbf{\hat x}_i, \ldots, \mathbf{\hat x}_N$ to the $\ell_0$ optimization problem:
%\begin{align}\label{minsum}
%\arg \min \sum_{i = 1}^N \|\mathbf{\hat x}_i\|_0 \ \ \text{s.t.} \ \ \mathbf{B}\mathbf{\hat x}_i = \mathbf{A}\mathbf{x}_i, \ \ \text{$i \in [N]$}, %\|\mathbf{A}\end{align}
%necessarily satisfies the recovery inequalities in \eqref{def1}. % \eqref{Cstable} and \eqref{b-PDa}.
%\end{corollary}
%
%\begin{proof}
%We derive a lower bound on the number of $k$-sparse $\mathbf{\hat x}_i$ and then apply Thm.~\ref{DeterministicUniquenessCorollary}. 
%First, note that $\|\mathbf{\hat x}_i\|_0 \neq 0$ for all $i$, since otherwise we would have $\mathbf{A}\mathbf{x}_i = \mathbf{B}\mathbf{\hat x}_i = \mathbf{0} \implies \mathbf{x}_i = \mathbf{0}$ by the spark condition, contradicting the general linear position of the $\mathbf{x}_i$. Let $n_p$ be the number of $\mathbf{\hat x}_i$ with $\|\mathbf{\hat x}_i\|_0 = p$.  %, so that $\sum_{p = 1}^{m} n_p = N$. 
%Since the $\mathbf{x}_i$ are all $k$-sparse, by \eqref{minsum} we have:
%\begin{align}
%k \sum_{p = 1}^{m} n_p = kN \geq \sum_{i=1}^N \|\mathbf{x}_i\|_0 \geq \sum_{i=1}^N \|\mathbf{\hat x}_i\|_0 = \sum_{p=1}^{m} p n_p.
%\end{align}
%%\[\implies \sum_{p = 1}^k (k-p)n_p \geq \sum_{p = k+1}^{m} (p-k) n_p.\]
%Hence,
%\begin{align}
%\sum_{p = k+1}^m n_p \leq \sum_{p = k+1}^m (p-k) n_p \leq \sum_{p = 1}^k (k-p)n_p \leq (k-1) \sum_{p = 1}^{k-1} n_p.
%\end{align}
%%\[ \implies \sum_{p = k+1}^m n_p \leq (k-1)^2 {\hat m \choose k-1} \]
%
%By the spark condition and general linear position of the $\mathbf{x}_i$, every $k$ vectors $\mathbf{Ax}_i$ span a $k$-dimensional space; hence, at most $k-1$ vectors $\mathbf{\hat x}_i$ can share a support of size $k-1$ or less, and the number of $(k-1)$-sparse vectors $\mathbf{\hat x}_i$ is then bounded from above by $(k-1) { m \choose k-1}$. It follows that at least $N - (k-1)^2 {m \choose k-1}$ of the vectors $\mathbf{\hat  x}_i$ are $k$-sparse, and that there must then be some $N' \subseteq [N]$ for which for each $S \in E$ there are $(k-1){m \choose k}+1$ of the vectors $\mathbf{x}_i$ with $i \in [N']$ supported on $S$ and for which $\mathbf{\hat x}_i$ is $k$-sparse for all $i \in [N']$.  The result now follows directly from Thm.~\ref{DeterministicUniquenessCorollary}. % \eqref{Cstable} and \eqref{b-PDa} follow.
%\end{proof}

@@

We  now show that $\pi$ is injective (and thus is a permutation if $\hat m = m$). Suppose that $\pi(i) = \pi(j) = \ell$ for some $i \neq j$ and $\ell$. Then, $\|c_{j}\mathbf{A}_{j} - \hat{c}_{j}\mathbf{B}_{\ell}\|_2 \leq \varepsilon$ and $\|c_{i}\mathbf{A}_{i} - \hat{c}_{i} \mathbf{B}_{\ell}\|_2  \leq \varepsilon$. Scaling and summing these inequalities by $|\hat{c}_{i}|$ and $|\hat{c}_{j}|$, respectively, and applying the triangle inequality, we obtain:
\begin{align*}%\label{contra}
(|\hat{c}_{i}| + |\hat{c}_{j}|) \varepsilon
&\geq\|\mathbf{A}(\hat{c}_{i}c_{j} \mathbf{e}_{j} - \hat{c}_{j}c_{i}\mathbf{e}_{i})\|_2 \nonumber \\ 
&\geq  \left( |\hat{c}_{i}| + |\hat{c}_{j}| \right) L_2(\mathbf{A}) \min_{\ell \in [m]} |c_\ell |,
\end{align*}
%
which contradicts the bound $\varepsilon < L_2(\mathbf{A})/C_1$. Hence, the map $\pi$ is injective. Setting $J = [m]$, $\hat J = \pi([m])$, and letting $P = \left( \mathbf{e}_{\pi(1)} \cdots \mathbf{e}_{\pi(m)}\right)$ and $D = \text{diag}(\frac{\hat{c}_1}{c_1},\ldots,\frac{\hat{c}_m}{c_m})$, we see that \eqref{1D} becomes, for all $j \in [m]$:
\begin{align*}%\label{k=1result}
\|\mathbf{A}_j - (\mathbf{B}_J\mathbf{PD})_j\|_2 
= \|\mathbf{A}_j - \frac{\hat{c}_j}{c_j}\mathbf{B}_{\pi(j)}\|_2 
\leq \frac{\varepsilon}{|c_j|} 
\leq C_1\varepsilon.
\end{align*}
\end{proof}

@@

If for some  $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ and map $\pi: E \mapsto {[\hat m] \choose k}$ we have:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon, \ \   \text{for all $S \in E$},
\end{align}
%
then $\hat m \geq m$ and, provided $p = \hat m - \ell(\hat m - m)$ is positive:
\begin{align}\label{MainLemmaBPD}
\|(\mathbf{A}_J - \mathbf{B}_{\hat J}\mathbf{PD})_j\|_2 \leq C_2 \varepsilon, \indent \text{for } j \in J.
\end{align}
for some $J \in {[m] \choose p}$, $\hat J \in {[\hat m] \choose p}$, permutation matrix $\mathbf{P}$, and invertible diagonal matrix $\mathbf{D}$.
\end{lemma}

@@

What we have shown here is that as an inverse problem to be approximated by a numerical algorithm, the sparse linear coding model  generally produces \textit{well-posed} optimizations.  This early concept of Hadamard \cite{Hadamard1902} can be paraphrased as the idea that good parameterized models should pose optimization problems with solutions that are \textit{stable}; that is, continuous in the data.  In other words, perturbing a dataset by a small amount should only slightly influence the identifiability or uniqueness of parameters in the model, rather than discontinuously affect the character of optimization involved.  In this regard, we demonstrate a linear relationship, \eqref{epsdel}, between noise in the data and the unique solution in  sparse linear coding, with explicit constants expressed in terms of original latent parameters. 
Nonetheless, it would be of practical utility to determine the best possible dependence of $\varepsilon$ on $\delta_1, \delta_2$ in Def.\ref{maindef} as well as the minimal requirements on the number and diversity of generating codes, and we hope that other researchers continue to improve and extend our results.
% , although they are too computationally expensive to be of practical value.  We hope .

@@

A technical difficulty in our main theorem is the absence of a spark condition assumption on alternate sparse data factorizations.
Although mathematically interesting that no such requirement is necessary, there are other reasons to seek out such a theoretical guarantee. For instance, it is difficult to ensure an algorithm maintain a dictionary satisfying \eqref{SparkCondition} at each learning stage; indeed, even certifying  a dictionary has this property is likely intractable given its NP-hardness \cite{tillmann2014computational}.
[*** TODO:  Mention relationship of $N$ being a polynomial in $(m,k)$ as giving an NP-hardness result using this Tillmann one ***]

In fact, though, uniqueness guarantees with minimal assumptions apply to all areas of data science and engineering that utilize learned sparse representations.  For example, several groups have utilized compressed sensing for signal processing tasks: MRI analysis \cite{lustig2008compressed},  image compression \cite{Duarte08}, and, more recently, the design of an ultrafast camera \cite{Gao14}. Given such effective uses of CS, it is only a matter of time before these systems incorporate sparse representational learning to encode and process data (e.g., in a device that learns structure from motion \cite{kong2016prior}).  In these cases, assurances such as those offered by our theorems certify that different devices (with different initialization, samples, etc.) are equivalent as soon as enough data originate from a statistically identical system.
%Data science:
%  
%in a common application of compressed sensing in engineering

%\textbf{Theoretical Neuroscience}.
Additionally, in the field of theoretical neuroscience, sparse dictionary learning and related methods have recovered characteristic components of natural images \cite{Olshausen96, hurri1996image, bell1997independent, van1998independent} and sounds \cite{bellsejnowski1996, smithlewicki2006, Carlson12}, reproducing response properties of cortical neurons. Our results suggest that this correspondence could be due to  sparse representations being ``universal'' in natural data, an early mathematical idea in neural theory \cite{pitts1947}. Furthermore, our guarantees justify the hypothesis of \cite{Coulter10, Isely10} that sparse codes passed through a communication bottleneck in the brain are recovered from random projections via (unsupervised) biologically plausible sparse dictionary learning (e.g., \cite{rehnsommer2007, rozell2007neurally, hu2014hebbian}).

@@

\begin{lemma}\label{NonEmptyLemma} 
Fix integers $\ell, m$ and $\hat m$. Suppose $E \subseteq 2^{[m]}$ is $\ell$-regular and satisfies the singleton intersection property. If there exists a map $\pi: E \to 2^{[\hat m]}$ such that $\sum_{S \in E} |S| = \sum_{S \in E} |\pi(S)|$ and:
\begin{align}\label{cond}
|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for all } F \subseteq E,
\end{align}
%
then $\hat m \geq m$ and the association $i \mapsto \cap \pi(F(i))$ defines an injective map from $J$ to $[\hat m]$ for some $J \subseteq [m]$ of size $\hat m - \ell(\hat m - m)$. In particular, if $\hat m = m$ then the map $\pi$ is induced by a permutation.
%$\pi(E)$ is $\ell$-regular and satisfies the singleton intersection property. In particular, $|\cap_{S \in F} S| = 1$ if and only if $|\cap_{S \in F} \pi(S)| = 1$. 
\end{lemma}
\begin{proof}
We first show that $\hat m \geq m$. Fix $T = \{(j, S): j \in \pi(S), S \in E\}$. Then $|T| = \sum_{S \in E} |\pi(S)| = \sum_{S \in E} |S| = \sum_{i \in [m]} \deg(i) = \ell m$ by regularity of $E$. If $\hat m < m$ then by pigeonholing the $m \ell \geq \ell (\hat m + 1) \geq [(\ell + 1) - 1] \hat m + 1$ elements of $T$ with respect to the set $[\hat m]$ of possible first indices, we see that there must exist at least $\ell + 1$ elements of $T$ having the same first index. But by \eqref{cond} there can be no more than $\ell$ elements of $T$ with a given first index, since $E$ is $\ell$-regular. Hence $\hat m \geq m$.
